---
title: "Analysis for RGC Dendrite Project (Chirp correlation)"
author: "Philipp Berens"
date: "July 31, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('ggplot2')

library("plyr")
library("mgcv")
library("hrbrthemes")
library("ggthemes")
library("itsadug")
library("dplyr")
library("xtable")


```

## Loading data


```{r load data}

data <- read.csv("../../data/processed/lchirp_corr.csv", header=TRUE)

data <- plyr::rename(data, c("lChirpCorr_distance"="roi_dist", "lChirpCorr_ExpDate"="exp_date"))

data <- plyr::rename(data, c("lChirpCorr_Angle"="angle", "lChirp_corr"="corr"))

data <- dplyr::select(data, c( "type", "roi_dist", "corr", "angle", "exp_date"))

data$type <- factor(data$type)
data$type <- revalue(data$type, c("0"="alpha transient", "1"="sustained", "2" = "mini alpha"))


# we filter the data to ROI pairs with distance <300 
# so that all cells are compared on equal grounds
# we use cell id as a random factor
ldata <- filter(data, data$roi_dist<300)
ldata <- filter(ldata, ldata$angle<180)
ldata$cell_id = factor(ldata$exp_date)


data <- read.csv("../../data/processed/chirp_corr.csv", header=TRUE)

data <- plyr::rename(data, c("ChirpCorr_distance"="roi_dist", "ChirpCorr_ExpDate"="exp_date"))

data <- plyr::rename(data, c("ChirpCorr_Angle"="angle", "Chirp_corr"="corr"))

data <- dplyr::select(data, c( "type", "roi_dist", "corr", "angle", "exp_date"))

data$type <- factor(data$type)
data$type <- revalue(data$type, c("0"="alpha transient", "1"="sustained", "2" = "mini alpha"))


# we filter the data to ROI pairs with distance <300 
# so that all cells are compared on equal grounds
# we use cell id as a random factor
gdata <- filter(data, data$roi_dist<300)
gdata <- filter(gdata, gdata$angle<180)
gdata$cell_id = factor(gdata$exp_date)

gdata$chirp = "global"
ldata$chirp = "local"

tdata = rbind(gdata,ldata)
tdata$chirp <- factor(tdata$chirp)


```

## Correlation vs. dendritic distance

Note that these first few plots here are mainly for getting an intuition about the data. Inferences from them should be drawn with care, since for each variable, I just average over all values of the other variable. For slices through the fitted model, see below. Also, I removed the SE bands, since they are not accurate as they do not take the repeated measures in individual cells into account.

### All cell types individually 
```{r chirp_distance, fig.width=7, fig.height=4, echo=FALSE, fig.path='figures/', dev=c('png', 'pdf')}
ggplot(tdata, aes(x=roi_dist,y=corr, color=chirp)) +
  geom_point(alpha=.1, shape=16, size=.5) + 
  geom_smooth(method = "gam", formula = y  ~  s(x), se=FALSE) + 
  facet_wrap(~type, ncol=3) + 
  ylim(c(0,1)) + 
  theme_minimal()+ 
  xlab(expression(paste("Dendritic distance [", mu, "m] "))) + 
  ylab("Correlation")
```

## Correlation vs. angular distance

###All cell types individually 
```{r chirp_angle, fig.width=7, fig.height=4, echo=FALSE, fig.path='figures/', dev=c('png', 'pdf')}
ggplot(tdata, aes(x=angle,y=corr, color=chirp, )) +
  geom_point(alpha=.1, shape=16, size=.5) + 
  geom_smooth(method = "gam", formula = y  ~  s(x), se=FALSE) + 
  facet_wrap(~type, ncol=3) + 
  ylim(c(0,1)) + 
  xlim(c(0,135)) +
  theme_minimal()+ 
  xlab("Angular distance [deg]") + 
  ylab("Correlation")

```

## Overlap index vs. distance and angle

### Gaussian GAM without random effects for cell id
```{r}
#m1 <- gam(corr ~ type + chirp +  te(roi_dist, angle, by=type, k=20), data = tdata)
#summary(m1)
#gamtabs(m1, type="HTML")
```


### Gaussian GAM with random effects for cell id

The first model is with type and chirp as additive effects:
```{r}
#m2 <- gam(corr ~ type + chirp + te(roi_dist, angle, by=type, k=20) +
#          s(cell_id, bs="re"), data = tdata)
#summary(m2)

```

The third model allows for an interaction between type and chirp:

```{r}
#m3 <- gam(corr ~ type * chirp + te(roi_dist, angle, by=type, k=30) +
#          s(cell_id, bs="re"), data = tdata)
#summary(m3)

```

The final model allows the smooth to depend on the interaction as well:

```{r}
tdata$typeXchirp = interaction(tdata$type, tdata$chirp)

m4 <- bam(corr ~ typeXchirp + te(roi_dist, angle, by=typeXchirp, k=50) +
          s(cell_id, bs="re"), data = tdata, discrete=TRUE)
summary(m4)


```



Model comparison between the two models using AIC:
```{r}
#AIC(m1)
#AIC(m2)
#AIC(m3)
AIC(m4)

```


Model comparison between the two models using ANOVA:

```{r}
#anova(m3,m4, test="F")
```

Because the AIC is more negative for m4 than any other one and the ANOVA yields a significant difference between the best two models, the random effects with smooths depending on the interaction model is better. Next, we want to check whether the assumptions of the model on the noise are warranted (Gaussianity):

```{r fig.width=4, fig.height=4}
#qq.gam(m4)
```

## Model visualization and analysis of the differences

We visualize the fitted surfaces and use the model to compute the differences between global and local chirps for each type and assess, where these are significantly different (assuming 99%-CIs). 

```{r chirp_distance_angle_talpha, fig.width=8.25, fig.height=4, echo=FALSE, fig.path='figures/', dev=c('png', 'pdf')}
par(mfrow=c(1,3), cex=1.1)

ng <- 60
dc = 1

zlimit = c(-1, 1) 
fvisgam(m4, view=c("roi_dist", "angle"), 
        cond=list(typeXchirp="alpha transient.local"),
        rm.ranef=TRUE, too.far=0.1, main="alpha transient, local", 
        xlab = expression(paste("Dendritic distance [", mu, "m] ")),
        ylab = "Angular distance (deg)",
        zlim=zlimit, n.grid = ng)

fvisgam(m4, view=c("roi_dist", "angle"), 
        cond=list(typeXchirp="alpha transient.global"),
        rm.ranef=TRUE, too.far=0.1, main="alpha transient, global", 
        xlab = expression(paste("Dendritic distance [", mu, "m] ")),
        ylab = "Angular distance (deg)",
        zlim=zlimit, n.grid = ng)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("alpha transient.local","alpha transient.global")),  
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = ng,zlim=c(-dc,dc), 
           print.summary = FALSE, hide.label = FALSE, se=2,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "diff", 
           add.color.legend=TRUE)
```


```{r chirp_distance_angle_minialpha, fig.width=8.25, fig.height=4, echo=FALSE, fig.path='figures/', dev=c('png', 'pdf')}
par(mfrow=c(1,3), cex=1.1)

fvisgam(m4, view=c("roi_dist", "angle"), 
        cond=list(typeXchirp="mini alpha.local"),
        rm.ranef=TRUE, too.far=0.1, main="mini alpha, local", 
        xlab = expression(paste("Dendritic distance [", mu, "m] ")),
        ylab = "Angular distance (deg)",
        zlim=zlimit, n.grid = ng)

fvisgam(m4, view=c("roi_dist", "angle"), 
        cond=list(typeXchirp="mini alpha.global"),
        rm.ranef=TRUE, too.far=0.1, main="mini alpha, global", 
        xlab = expression(paste("Dendritic distance [", mu, "m] ")),
        ylab = "Angular distance (deg)",
        zlim=zlimit, n.grid = ng)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("mini alpha.local","mini alpha.global")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = ng,zlim=c(-dc,dc), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "diff",
           add.color.legend=TRUE)


```

```{r chirp_distance_angle_sus, fig.width=8.25, fig.height=4, echo=FALSE, fig.path='figures/', dev=c('png', 'pdf')}
par(mfrow=c(1,3), cex=1.1)


fvisgam(m4, view=c("roi_dist", "angle"), 
        cond=list(typeXchirp="sustained.local"),
        rm.ranef=TRUE, too.far=0.1, main="sustained, local", 
        xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
        zlim=zlimit, n.grid = ng)
fvisgam(m4, view=c("roi_dist", "angle"), 
        cond=list(typeXchirp="sustained.global"),
        rm.ranef=TRUE, too.far=0.1, main="sustained, global", 
        xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
        zlim=zlimit, n.grid = ng)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("sustained.local","sustained.global")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = ng,zlim=c(-dc,dc), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "diff",
           add.color.legend=TRUE)



```




```{r chirp_distance_angle_types, fig.width=8.25, fig.height=8, echo=FALSE, fig.path='figures/', dev=c('png', 'pdf')}
par(mfrow=c(2,3), cex=1.1)

zlimit = c(-1, 1) 


plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("alpha transient.local","mini alpha.local")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = 60,zlim=c(-1,1), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "alpha - mini",
           add.color.legend=TRUE)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("alpha transient.local","sustained.local")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = 60,zlim=c(-1,1), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "alpha - sus",
           add.color.legend=FALSE)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("mini alpha.local","sustained.local")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = 60,zlim=c(-1,1), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "mini - sus",
           add.color.legend=FALSE)


plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("alpha transient.global","mini alpha.global")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = 60,zlim=c(-1,1), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "alpha - mini",
           add.color.legend=FALSE)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("alpha transient.global","sustained.global")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = 60,zlim=c(-1,1), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "alpha - sus",
           add.color.legend=FALSE)

plot_diff2(m4, view=c("roi_dist", "angle"), color="topo",
           comp=list(typeXchirp=c("mini alpha.global","sustained.global")),  se=2, 
           show.diff = TRUE, alpha.diff = .75, col.diff = 'white', 
           col='black',rm.ranef=TRUE, ci.col=c('gray45','gray80'), n.grid = 60,zlim=c(-1,1), 
           print.summary = FALSE, hide.label = FALSE,
           xlab = expression(paste("Dendritic distance [", mu, "m] ")),
           ylab = "Angular distance (deg)",
           main = "mini - sus",
           add.color.legend=FALSE)


```



